---
title: "Depletion calculations"
output: rmarkdown::github_document
---

# Overview

This script is intended to calculate fractional streamflow depletion (`Qf`) using the Glover model. 
`Qf` is depletion as a fraction of pumping rate. To obtain volumetric streamflow depletion (`Qs`), multiply `Qf` by the pumping rate (`Qw`). `Qs = Qf*Qw`

# Set up workspace

```{r echo = T, results = 'hide', message = F, warning = F}

## load packages
library(streamDepletr)
library(lubridate)
library(tidyverse)
library(sf)
library(patchwork)
options(dplyr.summarise.inform=F)   # suppress summarize info

# directory to keep stuff that is too big for GitHub
dir_big_files <- "C:/Users/samzipper/OneDrive - The University of Kansas/Research/StreamflowDepletion/DepletionMetrics_LapidesEtAl/data"

```

# Prep input data for model

```{r}

## times to test
n_yrs <- 50
times <- seq(1, 365*n_yrs, 1)

# for seasonal pumping, start/stop times
times_start <- seq(yday(ymd("2021-05-01")), yday(ymd("2021-05-01"))+365*n_yrs, 365)
times_stop <- seq(yday(ymd("2021-09-30")), yday(ymd("2021-09-30"))+365*n_yrs, 365)

## load gage shapefile
area_thres_km2 <- 10000  # only analyze gages < 10,000 km2
sf_gages <- 
  file.path("..", "Data", "gagesII_data_and_subsets", "gagesII.shp") %>% 
  st_read() %>% 
  mutate(GAGE_ID = as.numeric(STAID)) %>% 
  subset(DRAIN_SQKM <= area_thres_km2) %>% 
  subset(STAID != "15081497" &
           STAID != "15085100" &
           STAID != "15261000")  # no transmissivity data

# save final gages used for analysis
st_write(sf_gages, file.path("..", "Data", "depletion", "Gages_Depletion.gpkg"), append = F)

## load porosity from GLHYMPS: Porosity = value, GLHYMPS_Shape_Area = area of each porosity value within watershed
# since GLHYMPS gives porosity, may want to reduce to represent storativity since not all porosity
# is drainable. however, de Graaf et al. 2015 HESS and 2017 ADWR use these values directly for their models.
porosity <- 
  read_csv(file.path("..", "Data", "GLHYMPS_Porosity_by_gauges.csv")) %>% 
  subset(GAGE_ID %in% sf_gages$STAID) %>% 
  mutate(S = Porosity)  # could apply transformation if desired (i.e., S = 50% of porosity)

## there is a separate transmissivity file for each gage, load it within loop below

```

# Use Jasechko et al data to determine well-stream distance

```{r echo = T, message = F, warning = F}
## load well-stream distances
jsko_files <- list.files(file.path("..", "Data", "JasechkoEtAl_Fig2Data"), pattern = ".csv")
for (f in jsko_files){
  # compile all individual states into one data frame
  state <- gsub(".csv", "", f)
  jsko_state_data <- 
    file.path("..", "Data", "JasechkoEtAl_Fig2Data", f) %>% 
    read_csv() %>% 
    mutate(State = state)
  
  if (f == jsko_files[1]){
    jsko_all_states <- jsko_state_data
  } else {
    jsko_all_states <- bind_rows(jsko_all_states, jsko_state_data)
  }
}

# get median by state and overall
median_dist_m <- median(jsko_all_states$NearestFlowlineDistance_m)
jsko_all_states %>% 
  group_by(State) %>% 
  summarize(median_dist_m = median(NearestFlowlineDistance_m))

# plot distribution
ggplot(jsko_all_states, aes(x = NearestFlowlineDistance_m)) +
  geom_histogram(aes(fill = State), binwidth = 100) +
  theme_bw() 

ggplot(jsko_all_states, aes(x = State, y = NearestFlowlineDistance_m)) +
  geom_boxplot(aes(fill = State)) +
  geom_hline(yintercept = median_dist_m, color = "red") +
  theme_bw()

```

# Depletion calculation for each gage

```{r echo = T, message = F, warning = F}
## loop through all gages
n_iter <- 100   # number of iterations for each gage
for (usgs_id in sf_gages$STAID){
  # usgs_id <- sf_gages$STAID[1]
  
  ## build glover model inputs for each iteration sample
  # Well-stream distance [m]
  # this is the median well-stream distance from all wells in 12 states with data shared from Jasechko et al.
  # the median of all wells is 318 m. the lowest state median is KY (246 m) and the highest is TX (437 m).
  # could use the data to randomly sample from a distribution for uncertainty analysis.
  #   gage_d  <- median_dist_m # constant value everywhere
  set.seed(1)
  gage_d_sample <- sample(jsko_all_states$NearestFlowlineDistance_m, size = n_iter)
  
  # Storativity [-]
  porosity_gage <- subset(porosity, GAGE_ID == usgs_id)
  gage_S_sample <- sample(porosity_gage$S, size = n_iter, replace = T, prob = porosity_gage$GLHYMPS_Shape_Area)
  
  # Aquifer transmissivity [m2/d]
  df_Tr_gage <- 
    file.path("..", "Data", "transmissivity_data", paste0(usgs_id, ".csv")) %>% 
    read_csv(col_types = cols()) %>% 
    subset(complete.cases(.))
  names(df_Tr_gage) <- "Tr_m2d"
  
  if (dim(df_Tr_gage)[1] < n_iter){
    gage_Tr_sample <- sample(df_Tr_gage$Tr_m2d, size = n_iter, replace = T)
  } else {
    gage_Tr_sample <- sample(df_Tr_gage$Tr_m2d, size = n_iter, replace = F)
  }
  
  ## loop through iterations and calculate depletion
  for (n in 1:n_iter){
    ## get parameters for this sample
    gage_d <- round(gage_d_sample[n], 3)
    gage_S <- round(gage_S_sample[n], 3)
    gage_Tr <- gage_Tr_sample[n]
    
    ## calculate depletion
    # constant pumping
    Qf_constant <- glover(t = times, d = gage_d, S = gage_S, Tr = gage_Tr)
    # plot(times, Qf_constant)
    
    # seasonal pumping (may-sept)
    Qf_seasonal <- intermittent_pumping(t = times, starts = times_start, stops = times_stop, rates = 1, 
                                        method = "glover", d = gage_d, S = gage_S, Tr = gage_Tr)
    # plot(times, Qf_seasonal)
    
    ## summarize data
    daily_iter_depletion <- tibble(iter = n,
                                   day = times,
                                   year = rep(seq(1, n_yrs, 1), each = 365),
                                   DOY = rep(seq(1, 365, 1), times = n_yrs),
                                   Qf_constant = round(Qf_constant, 3),
                                   Qf_seasonal = round(Qf_seasonal, 3))
    
    ## calculate annual depletion
    annual_iter_depletion <- 
      daily_iter_depletion %>% 
      dplyr::group_by(iter, year) %>% 
      dplyr::summarize(Qf_constant_sum = sum(Qf_constant),
                       Qf_seasonal_sum = sum(Qf_seasonal))
    
    ## figure out years to dynamic equilibrium
    eq_thres <- 0.01  # 1% change in annual depletion volume
    prc_change_constant <- c(NA, diff(annual_iter_depletion$Qf_constant_sum))/annual_iter_depletion$Qf_constant_sum
    prc_change_seasonal <- c(NA, diff(annual_iter_depletion$Qf_seasonal_sum))/annual_iter_depletion$Qf_seasonal_sum
    eq_yrs_constant <- min(annual_iter_depletion$year[which(prc_change_constant < eq_thres)])
    eq_yrs_seasonal <- min(annual_iter_depletion$year[which(prc_change_seasonal < eq_thres)])
  
    ## summarize output data to save
    gage_iter_summary <- tibble(GAGE_ID = usgs_id,
                                iter = n,
                                dist_m = gage_d,
                                S = gage_S,
                                Tr_m2d = gage_Tr,
                                eq_yrs_constant = eq_yrs_constant,
                                eq_yrs_seasonal = eq_yrs_seasonal,
                                Qf_100yrs_constant = daily_iter_depletion$Qf_constant[length(times)])
          
    ## compile data frame
    if (n == 1){
      gage_summary <- gage_iter_summary
      daily_depletion <- daily_iter_depletion
      annual_depletion <- annual_iter_depletion
    } else {
      gage_summary <- bind_rows(gage_summary, gage_iter_summary)
      daily_depletion <- bind_rows(daily_depletion, daily_iter_depletion)
      annual_depletion <- bind_rows(annual_depletion, annual_iter_depletion)
    }
  }
  
  ## summarize to timeseries of daily median and stats
  daily_depletion_summary <-
    daily_depletion %>% 
    select(day, Qf_constant, Qf_seasonal) %>% 
    pivot_longer(-day, names_to = "pumping", values_to = "Qf") %>% 
    group_by(day, pumping) %>% 
    summarize(Qfmean = round(mean(Qf), 3), 
              Qfsd = round(sd(Qf), 3), 
              Qfmin = round(min(Qf), 3),
              Qf5 = round(quantile(Qf, probs = .05), 3),
              Qf25 = round(quantile(Qf, probs = .25), 3),
              Qf50 = round(quantile(Qf, probs = .5), 3),
              Qf75 = round(quantile(Qf, probs = .75), 3),
              Qf95 = round(quantile(Qf, probs = .95), 3),
              Qfmax = round(max(Qf), 3))
  daily_depletion_summary$pumping <- str_sub(daily_depletion_summary$pumping, start = 4)
  
  ## save output
  write_csv(daily_depletion_summary, file.path(dir_big_files, paste0("DepletionFraction_", usgs_id, "_DailySummary.csv")))
  write_csv(gage_summary, file.path("..", "Data", "depletion", paste0("DepletionFraction_", usgs_id, "_IterationSummary.csv")))
  
  ## status update
  print(paste0(which(sf_gages$STAID == usgs_id), " of ", length(sf_gages$STAID), " complete, ", Sys.time()))
}

```